### 2018-11-20 chap4.tex

%% This is an example first chapter.  You should put chapter/appendix that you
%% write into a separate file, and add a line \include{yourfilename} to
%% main.tex, where `yourfilename.tex' is the name of the chapter/appendix file.
%% You can process specific files by typing their names in at the
%% \files=
%% prompt when you run the file main.tex through LaTeX.

\chapter{Framework for rational donor selection in fecal microbiota transplant clinical trials}

The contents of this chapter were submitted with
Claire Duvallet, Caroline Zellmer, Pratik Panchal, Shrish Budree, Majdi Osman, and Eric J. Alm as authors.

The supplementary information is in Appendix B.
The supplementary tables are at the end of the chapter.

\includepdf[pages=-, pagecommand={\thispagestyle{plain}}, scale=0.95]{fmt/ms}