### 2018-11-27 abstract.tex
Edit abstract + first part of intro

%% The text of your abstract and nothing else (other than comments) goes here.
%% It will be single-spaced and the rest of the text that is supposed to go on
%% the abstract page will be generated by the abstractpage environment.

The human microbiome is essential for health and implicated in many diseases.
DNA sequencing has enabled the detailed characterization of human-associated microbial communities, leading to a rapid expansion in studies investigating the human microbiome and disease.
However, extracting clinically-relevant associations from microbiome datasets remains challenging because of the high dimensional nature of the data and variability across studies.
In this thesis, I describe three projects which overcome various analysis challenges to extract useful clinical insights from microbiome data.
In the first project, I present an analysis of lung, stomach, and oropharyngeal microbiomes.
I leverage data collected from multiple sites per patient to identify aspiration-associated changes in the relationships between aerodigestive communities which new approaches for treatment.
%These changes suggest new approaches for developing diagnostics and treatments for aspiration-related respiratory complications.
%I leverage data collected from multiple sites per patient to find aspiration-associated changes in the relationships between aerodigestive communities which suggests new targets for treatment and diagnosis.
In the second project, I perform a meta-analysis of case-control gut microbiome datasets with standard data processing and analysis methods.
I find consistent patterns characterizing disease-associated microbiome changes and a set of shared associations which could inform clinical treatment and therapeutic development approaches.
In the third project, I describe a framework for rational donor selection in fecal microbiota transplant clinical trials.
In this framework, knowledge derived from clinical and basic science research is used to inform which donor is selected for fecal transplants, increasing the likelihood of successful trials.
Together, these projects demonstrate a variety of approaches to mine the human microbiome for clinically-relevant insights and suggests multiple avenues forward for translating findings from microbiome data analyses into clinical impact.