### 2018-12-10 chap3.tex
Small changes

%% This is an example first chapter.  You should put chapter/appendix that you
%% write into a separate file, and add a line \include{yourfilename} to
%% main.tex, where `yourfilename.tex' is the name of the chapter/appendix file.
%% You can process specific files by typing their names in at the
%% \files=
%% prompt when you run the file main.tex through LaTeX.

\graphicspath{{meta-analysis/figures/}}

\chapter{Meta-analysis of gut microbiome studies identifies disease-specific and shared responses}

The contents of this chapter were published as:

Claire Duvallet, Sean M. Gibbons, Thomas Gurry, Rafael A. Irizarry, and Eric J. Alm. "Meta-analysis of gut microbiome studies identifies disease-specific and shared responses." \textit{Nature Communications}. (2017)

The figures are at the end of the chapter. The Supplementary Information is in Appendix B.

\clearpage
\input{meta-analysis/manuscript.tex}