### 2018-12-15 chap_24hr.tex

%% This is an example first chapter.  You should put chapter/appendix that you
%% write into a separate file, and add a line \include{yourfilename} to
%% main.tex, where `yourfilename.tex' is the name of the chapter/appendix file.
%% You can process specific files by typing their names in at the
%% \files=
%% prompt when you run the file main.tex through LaTeX.

%\graphicspath{{donor-selection/figures/}}

\chapter{Residential sewage provides a view into community-level human microbiome and metabolome}

\noindent
Mariana G Matus, Claire Duvallet, Melissa Soule, Noriko Endo, Newsha Ghaeli,  Ilana Brito, Carlo Ratti, Elizabeth B Kujawinski, Eric J Alm

\bigskip
\bigskip
\noindent
The contents of this chapter constitute a manuscript in preparation for submission.

\clearpage
%\input{donor-selection/ms.tex}