### 2018-12-15 chap_aspiration.tex

%% This is an example first chapter.  You should put chapter/appendix that you
%% write into a separate file, and add a line \include{yourfilename} to
%% main.tex, where `yourfilename.tex' is the name of the chapter/appendix file.
%% You can process specific files by typing their names in at the
%% \files=
%% prompt when you run the file main.tex through LaTeX.

\graphicspath{{aspiration/figures/}}

\chapter{Multi-site sampling reveals altered relationships in aerodigestive microbiomes of children with swallowing dysfunction}

Claire Duvallet, Kara Larson, Scott Snapper, Sonia Iosim, Ann Lee, Katherine Freer, Kara May, Eric J Alm, and Rachel Rosen.

\bigskip
\bigskip
\noindent
The contents of this chapter have been submitted for publication as: ``Aerodigestive sampling reveals altered microbial exchange between lung, oropharyngeal, and gastric microbiomes in children with impaired swallow function,'' in review as of December 2018.

\bigskip
\bigskip
\noindent
The tables and figures are at the end of the Chapter. The Supplementary Information is in Appendix A.

\clearpage
\input{aspiration/plosone_aspiration_main.tex}