### 2018-12-16 chap_donor-selection.tex

%% This is an example first chapter.  You should put chapter/appendix that you
%% write into a separate file, and add a line \include{yourfilename} to
%% main.tex, where `yourfilename.tex' is the name of the chapter/appendix file.
%% You can process specific files by typing their names in at the
%% \files=
%% prompt when you run the file main.tex through LaTeX.

\graphicspath{{donor-selection/figures/}}

\chapter{Increasing fecal microbiota transplant clinical trial successes by applying rational donor selection}\label{chap:donor-selection}

\noindent
Claire Duvallet, Caroline Zellmer, Pratik Panchal, Shrish Budree, Majdi Osman, and Eric J Alm.

\bigskip
\bigskip
\noindent
The contents of this chapter constitute a manuscript in preparation for submission.

\clearpage
\input{donor-selection/ms.tex}