### 2018-12-20 chap_meta-analysis.tex

%% This is an example first chapter.  You should put chapter/appendix that you
%% write into a separate file, and add a line \include{yourfilename} to
%% main.tex, where `yourfilename.tex' is the name of the chapter/appendix file.
%% You can process specific files by typing their names in at the
%% \files=
%% prompt when you run the file main.tex through LaTeX.

\graphicspath{{meta-analysis/figures/}}

\chapter{Meta-analysis of gut microbiome studies identifies disease-specific and shared responses}\label{chap:meta-analysis}

Claire Duvallet, Sean M Gibbons, Thomas Gurry, Rafael A Irizarry, and Eric J Alm.

\bigskip
\bigskip
\noindent
The contents of this chapter were published as: ``Meta-analysis of gut microbiome studies identifies disease-specific and shared responses'' in \textit{Nature Communications} (2017) volume 8, article number 1784.

\bigskip
\bigskip
\noindent
The figures are at the end of the chapter. The Supplementary Information is in Appendix \ref{app:meta-analysis}.

\clearpage
\input{meta-analysis/manuscript.tex}