### 2018-12-20 acknowledgments.tex
Move acknowledgments + make pared down versions of thesis

\section*{Acknowledgments}

Mentors:
Rafa - will always be a friend of the lab
Jim
Deb
Ilana - for being the mentor I didn't know I needed
Liz - teaching me everything I know about metabolomics and being so giving with time

Thomas - being my first introduction to computational work with patience and enthusiasm, enabling me to find my passion here
Mariana - drive and ability to make impossible things actually happen was an inspiration, expanded the scope of what a PhD student could do
Scott - being a role model, mentor, collaborator, and now close friend. Hope you remain all four forever. Also for encouraging me to start doing diversity and REFS
Sean Gibbons - modeling pure interest in science for the sake of science, encouraging me and being patient when I thought it was all fake, modeling success and being a joy to work with, constant sponsor of my work (opened a really important door into the world of open-source software)

Peers:
Nathaniel
Diversity co-conspirators: Scott and Manu, Bevin, and now my GSC crew for revitalizing me in the work (Nas, Josue, Danielle, German, Halston, and Bianca)

Funding:
NDSEG
Siebel

Friends, family, and frisbee
Carolyn and Megan
Andee and Shelby
Janyne for keeping me adventuring
sMITe
Felix (for GTD, 20% rule, and inspiration to version control everything)
ET (for putting us to shame by having a job and real life)
Parents (for supporting me, recognizing my passions better than I do, and listening to me talk about poop over many dinners)