### 2018-12-20 cover.tex

\title{Mining the human microbiome for clinical insight}

\author{Claire Marie No{\"e}lle Duvallet}

\prevdegrees{B.S., Columbia University (2013)}

\department{Department of Biological Engineering}

\degree{Doctor of Philosophy in Biological Engineering}

% Valid degree months are September,
% February, or June.  The default is June.
\degreemonth{June}
\degreeyear{2019}
\thesisdate{11 January 2019}

%% By default, the thesis will be copyrighted to MIT.  If you need to copyright
%% the thesis to yourself, just specify the `vi' documentclass option.  If for
%% some reason you want to exactly specify the copyright notice text, you can
%% use the \copyrightnoticetext command.
%\copyrightnoticetext{\copyright IBM, 1990.  Do not open till Xmas.}

% If there is more than one supervisor, use the \supervisor command
% once for each.
\supervisor{Eric J. Alm}{Professor}

% This is the department committee chairman, not the thesis committee
% chairman.  You should replace this with your Department's Committee
% Chairman.
\chairman{Forest White}{Chair of Graduate Program, Department of Biological Engineering}

% Make the titlepage based on the above information.  If you need
% something special and can't use the standard form, you can specify
% the exact text of the titlepage yourself.  Put it in a titlepage
% environment and leave blank lines where you want vertical space.
% The spaces will be adjusted to fill the entire page.  The dotted
% lines for the signatures are made with the \signature command.
\maketitle

% The abstractpage environment sets up everything on the page except
% the text itself.  The title and other header material are put at the
% top of the page, and the supervisors are listed at the bottom.  A
% new page is begun both before and after.  Of course, an abstract may
% be more than one page itself.  If you need more control over the
% format of the page, you can use the abstract environment, which puts
% the word "Abstract" at the beginning and single spaces its text.

% You can either \input (*not* \include) your abstract file, or you can put
% the text of the abstract directly between the \begin{abstractpage} and
% \end{abstractpage} commands.
\cleardoublepage
\setcounter{savepage}{\thepage}
\begin{abstractpage}
\input{abstract}
\end{abstractpage}

\cleardoublepage