### 2019-01-03 chap_24hr.tex
Update with latest versions

%% This is an example first chapter.  You should put chapter/appendix that you
%% write into a separate file, and add a line \include{yourfilename} to
%% main.tex, where `yourfilename.tex' is the name of the chapter/appendix file.
%% You can process specific files by typing their names in at the
%% \files=
%% prompt when you run the file main.tex through LaTeX.

% This graphicspath isn't working for some reason,
% need to manually write paths for each figure
\graphicspath{{24hr/figures/}}

\chapter{Residential sewage provides a view into community-level human microbiome and metabolome}\label{chap:24hr}

\noindent
Mariana Matus, Claire Duvallet, Melissa Kido Soule, Krista Longnecker, Noriko Endo, Newsha Ghaeli, Ilana Brito, Carlo Ratti, Elizabeth B Kujawinski, Eric J Alm

\bigskip
\bigskip
\noindent
The contents of this chapter constitute a manuscript in preparation for submission. The main figures are at the end of the chapter and the supplementary figures are in Appendix \ref{app:24hr}.

\clearpage
\input{24hr/24hr_ms.tex}